Detecting social information in a dense dataset of infants’ natural visual experience

### 2020-01-28 paper/cogsci/paper_jan28.Rmd
change datapaths in .Rmd to data_cogsci

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

@@

```{r}
## Read in data 
### Load preprocessed and filtered detections
load('../../data/openpose_detections_filtered_threekids.RData') #d_all
load('../../data/openpose_detections_filtered_cropped_allthreekids_jan28.RData') # d_cropped
```

@@

```{r}
### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
right_side_up_file = "../../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

@@

## load goal sample annotations
load('../../data/ketan_gold_sample.RData')

@@

```{r} 
### Load turk hand annotations wtih all bounding boxes (multiple dets per frame)
turk_annotations_seg = "../../data/turk_segmentations_hands_only_processed.csv" 
g_seg = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID) %>%
  mutate(center_x = left + width/2, center_y = top + height/2)

@@

```{r eval=FALSE}
hands_missed <- gold_sample %>%
  select(-vid_name) %>%
  rename(vid_name = vid_name_short) %>% 
  mutate(vid_name = as.factor(vid_name)) %>%
  left_join(child_adult_hand_annotations, by=(c("vid_name","frame"))) %>%
  filter(!is.na(full_image_path)) %>%
  filter(hand_eval_ketan == 'FN')

@@

```{r} 
gold_sample_no_child_hands <- gold_sample %>%
  left_join(child_adult_hand_annotations) %>%
  replace_na(list(child_hand_seg = FALSE)) %>% # replace NAs so they are counted
  filter(child_hand_seg==FALSE) %>%
  mutate(hand_eval_adults = evaluate_detector(hand_present_ketan, hand_openpose))

@@

We suspected that this is in part because children’s own hands were often in view of the camera and unconnected to a pose—a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s hands and adult’s hands. Overall, we found that almost half `r mean(hands_missed$child_hand_seg)` of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve to `r hand_performance_adults[3]` somewhat. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Figure 3.

@@

We thus re-analyzed the entire dataset while restricting our analysis to a smaller, upper portion of the frame comparable to the field of view used in @Fausey2016. To do so, we excluded hand detections that occurred in the bottom 40% of the frame, while retaining all hand detections that occurred in the top 60% of the frame. This coarse cropping of decreased the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)`percent to `r round(mean(d_cropped$hand_detected_cropped),2)` percent, but only decreased face detections from `r round(mean(d_all$face_openpose),2)` to `r round(mean(d_cropped$face_detected_cropped,2)`.  Within this modified field of view, we still observed, on average, more hand detections than face detections.

@@

```{r eval=FALSE}
summary_by_age_days_cropped_younger <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  filter(age_days < 730) %>% ## less than 24 months
  group_by(age_days,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected), num_detect = length(face_openpose)) %>%
  mutate(prop_faces = num_faces / num_detect, prop_hands = num_hands / num_detect)

summary_by_age_days_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  group_by(age_days,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected), num_detect = length(face_openpose)) %>%
  mutate(prop_faces = num_faces / num_detect, prop_hands = num_hands / num_detect)

@@

```{r eval=FALSE}
### Proportion faces vs. hands 
face_vs_hands_model_full_age <- lmer(prop_faces - prop_hands ~ scale(age_days) + scale(num_detect) + (age_days | child_id),  summary_by_age_days_cropped)

face_vs_hands_model_younger <- lmer(prop_faces - prop_hands ~ scale(age_days) + scale(num_detect) + (age_days | child_id),  summary_by_age_days_cropped_younger)

face_vs_hands_model_full_age <- glmer(cbind(num_faces,num_hands) ~ scale(age_days) + (1 | child_id), 
             summary_by_age_days_cropped, family = "binomial")
face_vs_hands_model_full_age_out <- summary(face_vs_hands_model_full_age)

# xtable(face_vs_hands_model_full_age_out$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. hands seen by infants.") 
```

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -0.26 & 0.09 & -2.83 & 0.005 \\ 
  scale(age\_months) & -0.15 & 0.00 & -225.40 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. hands seen by infants.} 
\end{table}

@@

### Variability by Location 
```{r}
load('../../data/child_adult_hand_annotations.RData')
load('../../data/saycam_metadata.RData')

>>>>

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
library(xtable)
theme_set(theme_few())
```

@@

```{r}
## Read in data 
### Load preprocessed and filtered detections
load('../../data_cogsci/openpose_detections_filtered_threekids_jan28.RData') #d_all
load('../../data_cogsci/openpose_detections_filtered_cropped_allthreekids_jan28.RData') # d_cropped
```

@@

```{r}
### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
right_side_up_file = "../../data_cogsci/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

@@

## load goal sample annotations
load('../../data_cogsci/ketan_gold_sample.RData')

@@

```{r} 
### Load turk hand annotations wtih all bounding boxes (multiple dets per frame)
turk_annotations_seg = "../../data_cogsci/turk_segmentations_hands_only_processed.csv" 
g_seg = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID) %>%
  mutate(center_x = left + width/2, center_y = top + height/2)

@@

```{r}
load('../../data_cogsci/child_adult_hand_annotations.RData')
# 
# hands_missed <- gold_sample %>%
#   select(-vid_name) %>%
#   rename(vid_name = vid_name_short) %>% 
#   mutate(vid_name = as.factor(vid_name)) %>%
#   left_join(child_adult_hand_annotations, by=(c("vid_name","frame"))) %>%
#   filter(!is.na(full_image_path)) %>%
#   filter(hand_eval_ketan == 'FN')

@@

```{r} 
load('../../data_cogsci/child_adult_hand_annotations.RData')

gold_sample_no_child_hands <- gold_sample %>%
  left_join(child_adult_hand_annotations) %>%
  replace_na(list(child_hand_seg = FALSE)) %>% # replace NAs so they are counted
  filter(child_hand_seg==FALSE) %>%
  mutate(hand_eval_adults = evaluate_detector(hand_present_ketan, hand_openpose))

@@

We suspected that this is in part because children’s own hands were often in view of the camera and unconnected to a pose—a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s hands and adult’s hands. Overall, we found that almost half  of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve to `r hand_performance_adults[3]` somewhat. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Figure 3.

@@

We thus re-analyzed the entire dataset while restricting our analysis to a smaller, upper portion of the frame comparable to the field of view used in @Fausey2016. To do so, we excluded hand detections that occurred in the bottom 40% of the frame, while retaining all hand detections that occurred in the top 60% of the frame. This coarse cropping of decreased the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)`percent to `r round(mean(d_cropped$hand_detected_cropped),2)` percent, but only decreased face detections from `r round(mean(d_all$face_openpose),2)` to `r round(mean(d_cropped$face_detected_cropped),2)`.  Within this modified field of view, we still observed, on average, more hand detections than face detections.

@@

```{r eval=FALSE}
summary_by_age_days_cropped_younger <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  filter(age_days < 730) %>% ## less than 24 months
  group_by(age_days,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected), num_detect = length(face_openpose)) %>%
  mutate(prop_faces = num_faces / num_detect, prop_hands = num_hands / num_detect) %>%
  mutate(num_no_detect_faces = num_detect - num_faces, num_no_detect_hands = num_detect - num_hands)

summary_by_age_days_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  group_by(age_days,child_id) %>%
  summarize(num_faces = sum(face_openpose), num_hands = sum(hand_detected), num_detect = length(face_openpose)) %>%
  mutate(num_no_detect_faces = num_detect - num_faces, num_no_detect_hands = num_detect - num_hands)

@@

```{r eval=FALSE}
### Proportion faces vs. hands 
faces_model_full_age <- glmer(cbind(num_faces,num_no_detect_faces) ~ scale(age_days) + (1 | child_id),  summary_by_age_days_cropped, family = "binomial")
# summary(faces_model_full_age)

hands_model_full_age <- glmer(cbind(num_hands,num_no_detect_hands) ~ scale((age_days)) + (1 | child_id),  summary_by_age_days_cropped, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
# summary(hands_model_full_age)

##
faces_model_younger <- glmer(cbind(num_faces,num_no_detect_faces) ~ scale((age_days)) + (1 | child_id),  summary_by_age_days_cropped_younger, family = "binomial",  control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

hands_model_younger <- glmer(cbind(num_hands,num_no_detect_hands) ~ scale((age_days)) + (1 | child_id),  summary_by_age_days_cropped, family = "binomial", ,  control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# faces_model_full_age_out <- summary(faces_model_full_age)
# xtable(faces_model_full_age_out$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces seen by infants.") 
```

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -1.55 & 0.03 & -52.47 & 0.000 \\ 
  scale(age\_days) & -0.13 & 0.00 & -261.57 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed model predicting the proportion of faces seen by infants.} 
\end{table}

@@

### Variability by Location 
```{r}
load('../../data_cogsci/saycam_metadata.RData')
